---
title: "What did the presidents say at their inauguration?"
output: html_notebook
---
# Step 0: check and install needed packages. Load the libraries and functions. 
```{r, message=FALSE, warning=FALSE}
packages.used=c(
  "dplyr",
  "readr",
  "DT",
  "tibble",
  "htmlwidgets",
  "xlsx",         # packages used for importing xlsx files
  "rJava",
  "xlsxjars",     
  "tm", # text mining package used for tf.countries-idf
  "wordcloud",
  "RColorBrewer",
  "tydytext",
  "maps", # for mapping
  "rworldmap",
  "sentimentr",
  "qdap",
  "RWeka"
  )

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# load packages
library("dplyr")
library("readr")
library("DT")
library("tibble")
library("htmlwidgets")
library("xlsx")
library("rJava")
library("xlsxjars")
library("tm")
library(tm)
library(wordcloud)
library(RColorBrewer)
library(dplyr)
library(tidytext)
library("sentimentr")
library("qdap")
library("RWeka")

```
# get inaugural speeches and metadata into dataframe
```{r}
inaug.list=read.xlsx(file.path("..","data","InaugurationInfo.xlsx",fsep = .Platform$file.sep),1)
speech.list=rbind(inaug.list)
speech.list$type=c(rep("inaugural", nrow(inaug.list)))
speech.list$id = c(seq(nrow(speech.list)))
speech.list$path=file.path("..","data","InauguralSpeeches",
                            paste0("inaug",speech.list$File,"-",speech.list$Term,".txt"),
                            fsep = .Platform$file.sep)
```
# get a single speech into a data frame so that we can figure out interesting objectives
```{r}
text_df <- data_frame(readLines(speech.list$path[1]))
text_df
```
# ome
```{r dependson = "text_df", R.options = list(dplyr.print_max = 10)}
text_df %>%
  unnest_tokens(c(text_df), txt, token = "sentences")
```


## Step 1.1: loop over text files to get speech text into dataframe
```{r, message=FALSE, warning=FALSE}
for(i in seq(nrow(speech.list))) {
  speech_text = paste(readLines(speech.list$path[i], n=-1, skipNul=TRUE, warn = FALSE), collapse = " ")
  speech.list$fulltext[i]=speech_text
}
```
## Step 1.3: count occurances
```{r}
ds = DirSource(directory = file.path("..","data","InauguralSpeeches",fsep = .Platform$file.sep),
               pattern = "*.txt",
               mode = "text")
control_list <- list(removePunctuation = TRUE,
                     removeWords = stopwords("en"),
                     tolower = TRUE,
                     removeNumbers = TRUE,
                     stripWhitespace = TRUE,
                     PlainTextDocument = TRUE)
speech.corpus = VCorpus(ds,readerControl = control_list)
speech.tdm = TermDocumentMatrix(speech.corpus, control = control_list)
```
# determine countries and continents mentioned in speeches
```{r}
require('maps')
us_cities = tolower(substr(us.cities$name,1,nchar(us.cities$name)-3)) # get city names
countries = tolower(unique(world.cities$country.etc)) # get country names
# append united kingdom

# russia was the soviet union
combined_speeches = PlainTextDocument(speech.corpus,language = "eng")
regions = c("europe","asia","africa","north america","south america","middle east","central america")

tf.countries = termFreq(combined_speeches,control = list(dictionary = countries))
tf.regions = termFreq(combined_speeches,control = list(dictionary = regions))

countries_mentioned <- names(as.list(tf.countries[tf.countries!=0]))

```
# summarize findings - which foreign country was mentioned the most times?
```{r}
term = "war"
tdm.tidy=tidy(speech.tdm)
tdm.overall=summarise(group_by(tdm.tidy, term), sum(count))
#hist(,plot=TRUE)
```
# Plotting foreign countries mentioned in Inaugural Addresses
```{r}
require(countrycode)
require(rworldmap)

theCountries <- countrycode(countries_mentioned,'country.name','iso3c')
# These are the ISO3 names of the countries you'd like to plot in red

menDF <- data.frame(country = theCountries,
  mentioned = TRUE)

menMap <- joinCountryData2Map(menDF, joinCode = "ISO3",
  nameJoinColumn = "country")
# This will join your menDF data.frame to the country map data

mapCountryData(menMap, nameColumnToPlot="mentioned", catMethod = "categorical",
  missingCountryCol = gray(.8), mapTitle = "Foreign Countries Mentioned in Inaugural Addresses", addLegend = FALSE)
# And this will plot it, with the trick that the color palette's first
# color is red
```
# Let's look at the context in which the countries were mentioned
```{r, message=FALSE, warning=FALSE}
sentence.list=NULL
for(i in 1:nrow(speech.list)){
  sentences=sent_detect(speech.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    sentence.list=rbind(sentence.list, 
                        cbind(speech.list[i,-ncol(speech.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
```